
# Introduction

```{r}
#| echo: false
#| include: false
library(MSEtool)
library(ggplot2)
rnorm(5)
RefOM <- readRDS('../objects/OM/Ref.om')
```

This document describes the trial specifications for the Management Strategy Evaluation (MSE) of South Atlantic albacore tuna (*Thunnus alalunga*), undertaken for the International Commission for the Conservation of Atlantic Tunas (ICCAT).

This is a living document and will be updated as the MSE process progresses.

## MSE Framework 

The MSE is built on the [openMSE](https://openmse.com/) framework (`MSEtool` v4+). All functions and code for reproducing the operating models, management procedures, and performance indicators are contained in the [`iccatSALB`](https://salb-mse.bluematterscience.com/pkg/) R package and its associated [GitHub repository](https://github.com/Blue-Matter/iccatSALB). The analysis itself is carried out through a set of scripts that call on these functions, which are also available in the repository alongside the package.

## Stock Assessment {#sec-stock-assessment}

The operating models are conditioned on the ICCAT South Atlantic albacore stock assessment, fitted using Stock Synthesis 3 (SS3) v3.30.23.1 [@methot2013], referred to here as the **S05** model run.

> TBD: link the assessment report once published

### Data

The S05 assessment is fit to catch and abundance index data from **1956--2024**, comprising:

- 5 catch fleets (`Fleet_01`--`Fleet_05`)
- 5 CPUE/abundance indices:
  - `Indx_CTP-LL_TB1` (Chinese Taipei longline, time block 1)
  - `Indx_CTP-LL_TB2` (Chinese Taipei longline, time block 2)
  - `Indx_JPN-LL_TB2` (Japan longline)
  - `Indx_BR_URY-LL` (Brazil/Uruguay longline)
  - `Indx_ZAF-BB` (South Africa baitboat)

### Model Structure

- **Model years:** 1956--2024 (69 years)
- **Areas:** 1 (single area, no movement)
- **Sexes:** 1 (combined-sex model)
- **Ages:** 0--15, with age 15 as the plus group

### Fixed Biological Parameters

The S05 model uses the following fixed/estimated biological parameter values:

| Parameter | Value |
|------------------------|------------------------|
| Natural mortality (*M*) | Lorenzen, reference *M* = 0.36 at reference age 11 |
| Growth model | von Bertalanffy |
| *L* at *A*min (age 0) | 23.17 cm |
| *L*∞ (*L* at *A*max) | 121.24 cm |
| von Bertalanffy *K* | 0.238 |
| CV (young) | 0.125 |
| CV (old) | 0.125 |
| Weight-length *a* | 1.3718e-05 |
| Weight-length *b* | 3.0793 |
| Maturity option | Age logistic |
| Age at 50% maturity | 5 |
| First mature age | 4 |
| Stock-recruitment functional form | Beverton-Holt |
| Steepness | 0.75 (fixed) |
| σR | 0.4 (fixed) |

```{r}
#| echo: false

## TODO: plots of growth, maturity, selectivity, etc
## RefOM@Stock[[1]]@Length@MeanAtAge[1,,1]

```

