# Forward Projections {#sec-forward-projections}

```{r}
#| include: false
library(MSEtool)
library(ggplot2)
rnorm(5)
OM <- readRDS("../objects/OM/Ref.om")
FirstManagementYear <- 2028
HistYears <- Years(OM,'H')
ProjYears <- Years(OM,'P')
```

This chapter describes the assumptions and settings for the projection years.

Each operating model has `r nSim(OM)` independent simulations. 

## Management Cycle and Data Lag

The MSE assumed that the first TAC generated by the CMPs will be set in
 `r FirstManagementYear`, with a `r Interval(OM)`-year
management cycle and and a `r DataLag(OM)`-year data lag .

This means that the TAC that will be implemented for the `r FirstManagementYear` fishing year 
will be calculated in `r FirstManagementYear-1` using data up to 
`r FirstManagementYear - DataLag(OM) - 1`, and then from there the TAC 
is updated every `r Interval(OM)` years (@tbl-mgmt-schedule).

```{r}
#| label: tbl-mgmt-schedule
#| tbl-cap: The management interval and data lag used in the SALB MSE. 
#| echo: false
#| results: asis
ManagementScheduleTable <- function(FirstYear, DataLag, Interval, 
                                    nYears = Interval * 2 + 1,
                                    print  = TRUE) {
  
  Years <- FirstYear + seq_len(nYears) - 1
  ManagementYears <- Years[seq(1, length(Years), by = Interval)]
  
  t <- Years - 1
  IsManagement <- Years %in% ManagementYears
  DataYear <- t - DataLag
  
  LastMgmtYear <- vapply(Years, \(y) max(ManagementYears[ManagementYears <= y]), numeric(1))
  
  TACDecision <- ifelse(
    IsManagement,
    sprintf("Calculated in %d, implemented %d", t, Years),
    sprintf("Carries forward %d's TAC", LastMgmtYear)
  )
  
  df <- data.frame(
    Year           = Years,
    t              = t,
    ManagementYear = ifelse(IsManagement, "Yes", "No"),
    TACDecision    = TACDecision,
    DataYearUsed   = ifelse(IsManagement, DataYear, "\u2014")
  )
  names(df) <- c("Implementation Year", "Calculation Year",
                 "Management Year?", "TAC decision", "DataYear used")
  
  header <- paste0("| ", paste(names(df), collapse = " | "), " |")
  sep    <- paste0("|", paste(rep("---", ncol(df)), collapse = "|"), "|")
  rows   <- apply(df, 1, \(r) paste0("| ", paste(r, collapse = " | "), " |"))
  
  if (print) {
    cat(c(header, sep, rows), sep = "\n")  
    return(invisible(df))
  }
  df
}

ManagementScheduleTable(FirstYear = 2028, DataLag = DataLag(OM), Interval = Interval(OM))
```


## Initial Projection Years

```{r}
#| include: false

implag     <- FirstManagementYear - max(HistYears) - 1
InitYears  <- ProjYears[seq_len(implag)]
DataYear1  <- FirstManagementYear - DataLag(OM) - 1

SRR    <- OM@Stock[[1]]@SRR
SDRec  <- round(SRR@SD[1], 2)
ACRec  <- round(SRR@AC[1], 2)
TruncSDRec <- SRR@TruncSD[1]
```

Following the stock assessment, the OMs have a historical period spanning from
`r min(HistYears)` to `r max(HistYears)`. Given that the TAC advice generated by
the CMPs will be implemented in `r FirstManagementYear`, the first `r implag` years of
the projection period (`r paste(InitYears, collapse = ", ")`) are, in effect, an
extension of the historical period rather than years managed by the CMPs. 

Where real catch data (and, for the indices used by index-based CMPs, real CPUE data)
become available for these years, they will be used directly in the OM. Where data are
not yet available, fleet-specific catches for these years will instead be simulated by
resampling from the mean and standard deviation of each fleet's catches over the
preceding four years.

This is necessary primarily because the population dynamics model needs a catch value for
every year in order to project the population forward in time. Additionally, the
CMPs require a complete data stream up to the data-lag year in order to calculate the first TAC. 

As shown in @tbl-mgmt-schedule, the `r FirstManagementYear` TAC is calculated in `r FirstManagementYear - 1` using data up to
`r DataYear1`. Because the historical period ends in `r max(HistYears)`, `r DataYear1` falls
within the projection period and will not yet be observed at the time this MSE is conditioned.
For now, the simulated `r DataYear1` catch and CPUE index values will be used to calculate the
`r FirstManagementYear` TAC; once real `r DataYear1` data become available, these can replace
the simulated values without requiring any change to the underlying MSE framework.

## Recruitment Deviations

Annual recruitment deviations are generated as autocorrelated, bias-corrected log-normal
deviates around the stock-recruit relationship. For a given simulation, the (log-scale)
deviation in year $y$ is calculated as:

$$
\varepsilon_y \sim \text{TruncNormal}(\mu, \sigma_R,\ \pm `r TruncSDRec`\sigma_R)
$$

$$
x_y = \rho\, x_{y-1} + \varepsilon_y \sqrt{1-\rho^2}
$$

$$
\text{RecDev}_y = \exp(x_y)
$$

where $\sigma_R$ is the recruitment deviation standard deviation, $\rho$ is the lag-1
autocorrelation, and $\mu = -0.5\sigma_R^2(1-\rho)/\sqrt{1-\rho^2}$ is a bias-correction
term so that $E[\text{RecDev}_y] = 1$ on the natural scale. Deviations are truncated at
$\pm `r TruncSDRec`\sigma_R$ (on the log scale) to avoid unrealistic extremes.


## Observation Error for Simulated Data

The data used by the CMPs (fleet-specific catches and standardised CPUE indices) are
simulated from the true OM-generated values with multiplicative observation error.

For catch, the observation error is conditioned on the historical fit between the OM's
simulated catch and the actual reported historical catch. In practice, the assessment model
is fit to the catch data essentially exactly, so the catch data provided to the CMPs 
exactly matches the catches in the OM.

```{r}
#| include: false
Hist       <- readRDS("../objects/Hist/Ref.hist")
StockName  <- names(Hist@OM@Obs)[1]
IndexNames <- grep("^Indx_", names(Hist@OM@Obs[[StockName]]), value = TRUE)

# last historical year with a non-NA observation error for a given index
LastHistYear <- function(ix) {
  Err      <- Hist@OM@Obs[[StockName]][[ix]]@Survey@Error
  HistErr  <- Err[, colnames(Err) %in% as.character(HistYears), drop = FALSE]
  HasData  <- !apply(HistErr, 2, function(x) all(is.na(x)))
  if (!any(HasData)) return(-Inf)
  max(as.numeric(colnames(HistErr)[HasData])) 
}

IndexObsSummary <- purrr::map_dfr(IndexNames, function(ix) {
  SurveyObs <- Hist@OM@Obs[[StockName]][[ix]]@Survey
  data.frame(
    Index        = ix,
    SD           = SurveyObs@Stats$SD[1],
    AC           = SurveyObs@Stats$AC[1],
    LastHistYear = LastHistYear(ix)
  )
})

# only indices still being produced in recent years continue to be generated
# in the projection period - historical indices that have since been
# discontinued are excluded from the example figure below
RecentIndexNames <- Filter(\(ix) LastHistYear(ix) >= max(HistYears) - 5, IndexNames)

IndexErrorDF <- purrr::map_dfr(RecentIndexNames, function(ix) {
  Array2DF(Hist@OM@Obs[[StockName]][[ix]]@Survey@Error) |>
    dplyr::mutate(Index = ix)
}) |>
  dplyr::filter(Sim <= 3, !is.na(Value))
```

For the CPUE indices, the
same historical-conditioning approach is used, but with a lag-1 autocorrelation additionally
estimated from the residuals between the observed and simulated index. Index observation
error therefore follows the same truncated, autocorrelated log-normal process described
above for recruitment deviations, continuing on from the last estimated historical residual.

@tbl-indexobserror shows the fitted SD and lag-1 autocorrelation for each index, together with
the last historical year in which the index was produced; all five indices are currently
assumed to scale proportionally with true abundance (i.e. no
hyperstability/hyperdepletion). 

These values are specific to the Reference OM shown here as an
example. The conditioned SD and autocorrelation will vary across the full set of
Reference and Robustness OMs, since each is conditioned on its own historical fit.

```{r}
#| label: tbl-indexobserror
#| echo: false
#| tbl-cap: "Conditioned CPUE index observation error for the Reference OM (as an example): SD of the log-scale residuals, lag-1 autocorrelation, and the last historical year each index was produced."
knitr::kable(
  IndexObsSummary,
  digits = 3,
  col.names = c('Index', 'SD (log-scale)', 'Lag-1 AC', 'Last historical year')
)
```

@fig-indexobserror shows example realisations of the index observation error multiplier for
three simulations, spanning the historical and projection periods, again for the Reference OM
only. 

It illustrates that the autocorrelated observation error applied in the projections has
the same statistical properties as estimated from the historical period, with no discontinuity
at the start of the projection period. Historical indices that are no longer produced are excluded, since
these will not be generated or used in the projection period.

```{r}
#| label: fig-indexobserror
#| echo: false
#| fig-cap: "Example CPUE index observation error multipliers for three simulations (Reference OM only), spanning the historical and projection periods. The dashed line marks the end of the historical period. Discontinued indices are excluded (see text)."
#| fig-width: 8
#| fig-height: 12
IndexErrorDF$Sim <- as.factor(IndexErrorDF$Sim)

IndexErrorDF <- IndexErrorDF |>
  dplyr::mutate(Period = ifelse(Year <= max(HistYears), 'Historical', 'Projection'))

# duplicate the boundary year into the projection period so the two segments
# join up visually, rather than leaving a gap where the linetype switches
BoundaryDF <- IndexErrorDF |>
  dplyr::filter(Year == max(HistYears)) |>
  dplyr::mutate(Period = 'Projection')

PlotDF <- dplyr::bind_rows(IndexErrorDF, BoundaryDF)
PlotDF$Index <- factor(PlotDF$Index, levels = unique(PlotDF$Index),ordered = TRUE)

ggplot(PlotDF, aes(x = Year, y = Value, colour = Sim, linetype = Period,
                   group = interaction(Sim, Period))) +
  geom_line() +
  facet_wrap( ~ Index, ncol = 1) +
  scale_linetype_manual(values = c(Historical = 'dashed', Projection = 'solid')) +
  labs(x = 'Year', y = 'Index observation error multiplier',
      colour = 'Simulation', linetype = 'Period') +
  theme_bw() +
  geom_hline(yintercept = 1, linetype = 3, color = 'darkgray') +
  theme(legend.position = 'bottom')
```



