---
suppress-bibliography: true
---

# Forward Projections {#sec-forward-projections}

```{r}
#| include: false
library(MSEtool)
library(ggplot2)
rnorm(5)
OM <- readRDS("../objects/OM/Ref.om")
FirstManagementYear <- 2028
HistYears <- Years(OM,'H')
ProjYears <- Years(OM,'P')
```

Each operating model has `r nSim(OM)` independent simulations with stochastic samples 
of the log-normally distributed recruitment deviations and observation error for
the indices of abundance (see below).

## Management Cycle and Data Lag

The MSE assumes that the first TAC generated by the CMPs will be set in
 `r FirstManagementYear`, with a `r Interval(OM)`-year
management cycle and a `r DataLag(OM)`-year data lag.

This means that the TAC that will be implemented for the `r FirstManagementYear` fishing year 
will be calculated in `r FirstManagementYear-1` using data up to 
`r FirstManagementYear - DataLag(OM) - 1`, and then from there the TAC 
is updated every `r Interval(OM)` years (@tbl-mgmt-schedule).

```{r}
#| label: tbl-mgmt-schedule
#| tbl-cap: The management interval and data lag used in the SALB MSE. 
#| echo: false
#| results: asis
MSEtool::ManagementScheduleTable(FirstYear = 2028, DataLag = DataLag(OM), Interval = Interval(OM))
```

## Initial Projection Years

```{r}
#| include: false

implag     <- FirstManagementYear - max(HistYears) - 1
InitYears  <- ProjYears[seq_len(implag)]
DataYear1  <- FirstManagementYear - DataLag(OM) - 1

SRR    <- OM@Stock[[1]]@SRR
SDRec  <- round(SRR@SD[1], 2)
ACRec  <- round(SRR@AC[1], 2)
TruncSDRec <- SRR@TruncSD[1]
```

Following the stock assessment, the OMs have a historical period spanning from
`r min(HistYears)` to `r max(HistYears)`. Given that the TAC advice generated by
the CMPs will be implemented in `r FirstManagementYear`, the first `r implag` years of
the projection period (`r paste(range(InitYears), collapse = " -- ")`) are, in effect, an
extension of the historical period rather than years managed by the CMPs. 

When real catch data and real CPUE data become available for these years, 
they will be used directly in the OM. While these data are not yet available, 
fleet-specific catches for these years are instead simulated.

For each fleet, an independent catch value is drawn for every year without real data from a
lognormal distribution with a natural-scale mean and SD set to that fleet's own mean
and SD over the preceding four historical years (@tbl-interim-advice). This is necessary 
because the population dynamics model needs a catch value for every year in order
to project the population forward in time. Additionally, the CMPs require a 
complete data stream up to the data-lag year in order to calculate the first TAC.

As shown in @tbl-mgmt-schedule, the `r FirstManagementYear` TAC is calculated in `r FirstManagementYear - 1` using data up to
`r DataYear1`. Because the historical period ends in `r max(HistYears)`, `r DataYear1` falls
within the projection period and will not yet be observed at the time the MSE was run.
For now, the simulated `r DataYear1` catch and CPUE index values will be used by the CMPs 
to calculate the `r FirstManagementYear` TAC. If real `r DataYear1` data become available, these can replace
the simulated values without requiring any change to the underlying MSE framework (although the CMPs may need to be re-tuned; see @sec-tuning-methodology).

```{r}
#| include: false
InterimAdviceDF <- MSEtool::InterimAdvice(OM)
InterimAdviceSummary <- InterimAdviceDF |>
  dplyr::distinct(Fleet, Mean, SD) |>
  dplyr::mutate(Fleet = as.character(Fleet))
InterimYearRange <- range(InterimAdviceDF$Year)
```

```{r}
#| label: tbl-interim-advice
#| echo: false
#| tbl-cap: !expr "paste0('The fleet-specific mean and SD (natural scale, metric tons) of the lognormal distribution that the interim catches for the ', InterimYearRange[1], '--', InterimYearRange[2], ' projection years are sampled from.')"
knitr::kable(
  InterimAdviceSummary,
  digits = 1,
  col.names = c('Fleet', 'Mean', 'SD')
)
```

## Recruitment Deviations

Annual recruitment deviations are generated as autocorrelated, bias-corrected log-normal
deviates around the stock-recruit relationship. For a given simulation, the (log-scale)
deviation in year $y$ is calculated as:

$$
\varepsilon_y \sim \text{TruncNormal}(\mu, \sigma_R,\ \pm `r TruncSDRec`\sigma_R)
$$

$$
x_y = \rho\, x_{y-1} + \varepsilon_y \sqrt{1-\rho^2}
$$

$$
\text{RecDev}_y = \exp(x_y)
$$

where $\sigma_R$ is the recruitment deviation standard deviation, $\rho$ is the lag-1
autocorrelation, and $\mu = -0.5\sigma_R^2(1-\rho)/\sqrt{1-\rho^2}$ is a bias-correction
term so that $E[\text{RecDev}_y] = 1$ on the natural scale. Deviations are truncated at
$\pm `r TruncSDRec`\sigma_R$ (on the log scale) to avoid unrealistic extremes.

```{r}
#| include: false
SDProjParam <- OM@Stock[[1]]@SRR@SD[1]
ACProjParam <- OM@Stock[[1]]@SRR@AC[1]
```

$\sigma_R$ and $\rho$ are set to the standard deviation and lag-1 autocorrelation 
of each OM's estimated historical recruitment deviations so that the process used 
to generate each projection year's deviations is calibrated to what was estimated historically (e.g., $\sigma_R=$ `r round(SDProjParam, 2)`,
$\rho=$ `r round(ACProjParam, 2)` for the Reference OM, the central OM at the 50th percentile of
the Growth and Natural Mortality uncertainty grid; @fig-recdev-deviations).

```{r}
#| label: fig-recdev-deviations
#| echo: false
#| fig-cap: "Recruitment deviations (log scale) across all simulations of the Reference OM showing the median (solid line) and the 5th and 95th percentiles (shading)."
#| fig-width: 9
#| fig-height: 4.5
MSEtool::PlotRecDevs(OM)
```

## Observation Error for Simulated Data

The data used by the CMPs (fleet-specific catches and standardised CPUE indices) are
simulated from the true OM-generated values with multiplicative observation error.

For catch, the observation error is conditioned on the historical fit between the OM's
simulated catch and the actual reported historical catch. In practice, the assessment model
is fit to the catch data essentially exactly, so the catch data provided to the CMPs 
exactly matches the catches in the OM.

```{r index_error}
#| include: false
Hist       <- readRDS("../objects/Hist/Ref.hist")
StockName  <- names(Hist@OM@Obs)[1]
IndexNames <- grep("^Indx_", names(Hist@OM@Obs[[StockName]]), value = TRUE)

# last historical year with a non-NA observation error for a given index
LastHistYear <- function(ix) {
  Err      <- Hist@OM@Obs[[StockName]][[ix]]@Survey@Error
  HistErr  <- Err[, colnames(Err) %in% as.character(HistYears), drop = FALSE]
  HasData  <- !apply(HistErr, 2, function(x) all(is.na(x)))
  if (!any(HasData)) return(-Inf)
  max(as.numeric(colnames(HistErr)[HasData])) 
}

IndexObsSummary <- purrr::map_dfr(IndexNames, function(ix) {
  SurveyObs <- Hist@OM@Obs[[StockName]][[ix]]@Survey
  data.frame(
    Index        = ix,
    SD           = SurveyObs@Stats$SD[1],
    AC           = SurveyObs@Stats$AC[1],
    LastHistYear = LastHistYear(ix)
  ) 
})

RecentIndexNames <- Filter(\(ix) LastHistYear(ix) >= max(HistYears) - 5, IndexNames)

IndexErrorDF <- purrr::map_dfr(RecentIndexNames, function(ix) {
  Array2DF(Hist@OM@Obs[[StockName]][[ix]]@Survey@Error) |>
    dplyr::mutate(Index = ix)
}) |>
  dplyr::filter(Sim <= 3, !is.na(Value))
```

For the CPUE indices, the same historical-conditioning approach is used,
but with a lag-1 autocorrelation additionally estimated from the residuals between
the observed and simulated index. Index observation error therefore follows the
same truncated, autocorrelated log-normal process described above for recruitment
deviations, continuing on from the last estimated historical residual.

@tbl-indexobserror shows, for the Reference OM, the fitted SD and lag-1 autocorrelation for each index that
is available in the terminal year. The conditioned SD and autocorrelation will vary across the full set of
Reference and Robustness OMs, since each is conditioned on its own historical fit.
All indices are assumed to scale proportionally with true abundance (i.e. no
hyperstability/hyperdepletion).


```{r}
#| label: tbl-indexobserror
#| echo: false
#| tbl-cap: "Conditioned CPUE index observation error for the Reference OM showing the SD of the log-scale residuals and the lag-1 autocorrelation for the three indices that are simulated in the projection period."
knitr::kable(
  IndexObsSummary |> 
    dplyr::filter(LastHistYear == max(LastHistYear)) |> 
    dplyr::select(Index, SD, AC),
  digits = 3,
  col.names = c('Index', 'SD (log-scale)', 'Lag-1 AC')
)
```

@fig-indexobserror shows example realisations of the index observation error multiplier for
three simulations, again using the Reference OM as an example. It illustrates that the autocorrelated observation error applied in the projections has
the same statistical properties as estimated from the historical period.

@fig-survey-vs-pop shows, using the Reference OM as an example, the comparison of the indices of abundance
in the historical and projection years against their corresponding vulnerable population abundance (biomass or numbers depending on each index's units).

```{r}
#| label: fig-indexobserror
#| echo: false
#| fig-cap: "Example CPUE index observation error multipliers for three simulations from the Reference OM."
#| fig-width: 8
#| fig-height: 12
IndexErrorDF$Sim <- as.factor(IndexErrorDF$Sim)

IndexErrorDF <- IndexErrorDF |>
  dplyr::mutate(Period = ifelse(Year <= max(HistYears), 'Historical', 'Projection'))

BoundaryDF <- IndexErrorDF |>
  dplyr::filter(Year == max(HistYears)) |>
  dplyr::mutate(Period = 'Projection')

PlotDF <- dplyr::bind_rows(IndexErrorDF, BoundaryDF)
PlotDF$Index <- factor(PlotDF$Index, levels = unique(PlotDF$Index),ordered = TRUE)

ggplot(PlotDF, aes(x = Year, y = Value, colour = Sim, linetype = Period,
                   group = interaction(Sim, Period))) +
  geom_hline(yintercept = 1, linetype = 2, color = 'black') +
  geom_line() +
  facet_wrap( ~ Index, ncol = 1) +
  scale_linetype_manual(values = c(Historical = 'dashed', Projection = 'solid')) +
  labs(x = 'Year', y = 'Index observation error multiplier',
      colour = 'Simulation', linetype = 'Period') +
  theme_bw() +
  expand_limits(y = 0) +
  theme(legend.position = 'bottom')
```



```{r}
#| include: false
ExMSE <- readRDS('../objects/MSE/Reference/IndexRateAll3/G_50-M_50.mse') 
```

```{r}
#| label: fig-survey-vs-pop
#| echo: false
#| fig-width: 9
#| fig-height: 12
#| fig-cap: "An example of a single simulation of the a) Chinese Taipei longline, b) Brazil-Uruguay longline, and c) South Africa baitboat indices plotted against the corresponding selectivity-weighted vulnerable biomass described in the operating model. The vertical dashed line indicates the end of the historical period (2024)."
p1 <- MSEtool::PlotIndexFit(ExMSE, type = 'Survey', Fleets = 'Indx_CTP-LL_TB2', probs = 0, nsim = 1) +
  ggplot2::geom_vline(xintercept = 2024, linetype = 2, color = 'darkgray') 
p2 <- MSEtool::PlotIndexFit(ExMSE, type = 'Survey', Fleets = 'Indx_BR_URY-LL', probs = 0, nsim = 1) +
  ggplot2::geom_vline(xintercept = 2024, linetype = 2, color = 'darkgray')   
p3 <- MSEtool::PlotIndexFit(ExMSE, type = 'Survey', Fleets = 'Indx_ZAF-BB', probs = 0, nsim = 1) +
  ggplot2::geom_vline(xintercept = 2024, linetype = 2, color = 'darkgray') 

patchwork::wrap_plots(p1, p2, p3, ncol = 1, guides = 'collect') +
  patchwork::plot_annotation(tag_levels = 'a', tag_suffix = ')')

```

