---
suppress-bibliography: true
---

# Candidate Management Procedures {#sec-cmps}

```{r}
#| include: false
library(MSEtool)
library(ggplot2)
library(iccatSALB)
source('../analysis/04-Define-CMPs.R')

FinalMP_Families <- setdiff(names(MP_Groups), c('Reference', 'MultiIndexAgreement'))
FinalMP_Groups <- MP_Groups[FinalMP_Families]
FinalMP_Groups$MCC <- 'MCC_98'

slick <- readRDS('../objects/slick.rds')
MPMeta <- data.frame(Code = Slick::Code(slick@MPs), Label = Slick::Label(slick@MPs))
mp_label <- function(code) MPMeta$Label[match(code, MPMeta$Code)]

hcr_curve <- function(ControlPointsIndex, ControlPointsRate, RampType = 'linear', xmax = 1.8, n = 400) {
  x <- seq(0, xmax, length.out = n)
  y <- vapply(x, function(xi) {
    MSEtool::HockeyStickHCR(TrialRate = 1, Est = xi, Ref = 1,
                             ControlPointsIndex = ControlPointsIndex,
                             ControlPointsRate = ControlPointsRate,
                             RampType = RampType)
  }, numeric(1))
  data.frame(x = x, y = y)
}

plot_hcr <- function(ControlPointsIndex, ControlPointsRate, RampType = 'linear',
                      xlab, ylab = 'Fraction of trial rate applied') {
  ggplot(hcr_curve(ControlPointsIndex, ControlPointsRate, RampType), aes(x, y)) +
    geom_line(linewidth = 0.9, color = '#0072B2') +
    expand_limits(x = 0, y = c(0, 1.05)) +
    labs(x = xlab, y = ylab) +
    theme_bw()
}

pct_cap <- function(dd, du) {
  if (isTRUE(all.equal(dd[2], du[2]))) {
    sprintf('%.0f%%', 100 * dd[2])
  } else {
    sprintf('%.0f%% down / %.0f%% up', 100 * dd[2], 100 * du[2])
  }
}

pct_range_of <- function(variant_name) {
  gen <- get(variant_name)
  pct_cap(eval(formals(gen)$DeltaDown), eval(formals(gen)$DeltaUp))
}
```

The MSE currently includes three model-free CMP families:

1. Index Rate
2. Mostly Constant Catch
3. Ensemble

There are two variants of the Index Rate and Ensemble CMPs, making a total five CMPs included in
the analysis (@tbl-cmp-summary). Each CMP is described in full in the sections below.

Model-based CMPs using a surplus production assessment model (SPiCT; @pedersenStochasticSurplusProduction2017) were
thoroughly explored but were not included in the final analysis due to the consistent poor performance relative to the
model-free CMPs.

## Description of CMPs 

```{r}
#| label: tbl-cmp-summary
#| echo: false
#| tbl-cap: "The Code and Label, tuning parameter values, and the maximum per-cycle TAC change (where applicable) for the five CMPs variants."
CMPCodes <- c('MCC', 'IR', 'IR_20', 'Ens', 'Ens_20')
CMPNames <- c(FinalMP_Groups$MCC, FinalMP_Groups$IndexRate, FinalMP_Groups$IndexRate_MC20,
              FinalMP_Groups$Ensemble, FinalMP_Groups$EnsembleIR20)

ensemble_ir_cap <- function(variant_name) {
  gen <- get(variant_name)
  paste(pct_cap(eval(formals(gen)$IR_DeltaDown), eval(formals(gen)$IR_DeltaUp)), '(IR)')
}

ensemble_note <- function(variant_name) {
  gen <- get(variant_name)
  sprintf('MCC tunepar %s, IR tunepar %s',
          eval(formals(gen)$MCC_tunepar), eval(formals(gen)$IR_tunepar))
}

data.frame(
  Code = CMPCodes,
  Label = mp_label(CMPCodes),
  tunepar = unname(vapply(CMPNames, function(n) eval(formals(get(n))$tunepar), numeric(1))),
  `Max change / cycle` = c(
    '--',
    pct_range_of(FinalMP_Groups$IndexRate),
    pct_range_of(FinalMP_Groups$IndexRate_MC20),
    ensemble_ir_cap(FinalMP_Groups$Ensemble),
    ensemble_ir_cap(FinalMP_Groups$EnsembleIR20)
  ),
  Notes = c(
    '--',
    '--',
    '--',
    ensemble_note(FinalMP_Groups$Ensemble),
    ensemble_note(FinalMP_Groups$EnsembleIR20)
  ),
  check.names = FALSE
) |> knitr::kable(row.names = FALSE, align = c('l', 'l', 'c', 'c', 'l'))
```

### Index Rate {#sec-cmps-indexrate}

The Index Rate CMPs use the Chinese Taipei longline, Brazil-Uruguay longline, and South Africa
baitboat indices, together with catch data from the last historical year, to set each management
cycle's trial catch limit in five steps:

1. Each index's current status is projected forward using its own short-term trend, and the three
   indices are combined (equally weighted) into a single current index value, $I_{\text{cur}}$,
   and a single historical reference level, $I_{\text{ref}}$.
2. Each index is separately calibrated into its own catch-per-index rate: the total catch across
   all fleets in the last historical year, divided by that index's own value in the calibration
   period.
3. A single ramped response, based on how $I_{\text{cur}}$ compares with $I_{\text{ref}}$, throttles
   all three indices' rates by the same fraction.
4. Each index's most recent value, multiplied by its own throttled rate, gives that index's own
   implied catch limit; the trial catch limit is the equally-weighted average of the three.
5. Only a fraction of the move from the previous cycle's catch limit to the trial catch limit is
   actually taken (the `Responsiveness` parameter), and the resulting change is then optionally further capped.

These steps are described in more detail below.

Each index's `Status` is simply its own most recent-year value. `Status` is then projected
forward to account for any short-term trend: a linear trend is fitted by least-squares regression to
the log of the index over the preceding $n_T$ years, and `Status` is extrapolated forward $h$ years
at that fitted rate:
  
$$
I_{\text{proj}} = \text{Status} \times e^{\,\text{Slope} \times h}
$$
where Slope is the fitted trend (in log-index units per year), $n_T$ =
`r eval(formals(IndexRate_MP)$TrendYears)` years is the trend-fitting window, and $h$ =
`r eval(formals(IndexRate_MP)$TrendHorizon)` years is how far forward the status is projected.

This calculation is done separately for each of the three indices, giving each index its own
projected value, $I_{\text{proj}}$ (@fig-indexrate-trend). These three $I_{\text{proj}}$ values are then combined into a single
current index, $I_{\text{cur}}$, as an equally-weighted average across the three indices. Each
index's own historical reference level, the average level over 2007--2010,  is combined in the
same mannger, using the same equal weighting, to give the combined reference level, $I_{\text{ref}}$.

```{r}
#| label: fig-indexrate-trend
#| warning: false
#| echo: false
#| fig-width: 9
#| fig-height: 4
#| fig-cap: "The status-projection step applied to all three indices used by the Index Rate CMPs showing the historical index (grey), the trend fitted over the preceding years used for the fit (solid blue) extended forward (dashed blue), the most recent year's value (filled point, Status), and the resulting projected value (open point, $I_{\\text{proj}}$)."
Hist       <- readRDS('../objects/Hist/Ref.hist')
IndexValue <- Hist@Data[[1]][[1]]@Survey@Value
IndexNames <- eval(formals(get(FinalMP_Groups$IndexRate))$Indices)

TrendYears   <- eval(formals(IndexRate_MP)$TrendYears)
TrendHorizon <- eval(formals(IndexRate_MP)$TrendHorizon)
RecentYears  <- eval(formals(IndexRate_MP)$RecentYears)
RefYears     <- 2007:2010

Year  <- as.numeric(rownames(IndexValue))
nYear <- length(Year)

SeriesList <- list(); TrendList <- list(); PointList <- list()
IProjVec <- numeric(0); IRefVec <- numeric(0)

for (nm in IndexNames) {
  Value <- IndexValue[, nm]

  FitRows  <- seq(max(1, nYear - TrendYears + 1), nYear)
  FitYear  <- Year[FitRows]
  FitValue <- Value[FitRows]
  Fit      <- lm(log(FitValue) ~ FitYear)
  Slope    <- unname(coef(Fit)[2])

  Status   <- mean(utils::tail(Value, RecentYears))
  ProjYear <- max(Year) + TrendHorizon
  IProj    <- Status * exp(Slope * TrendHorizon)
  IRef     <- mean(Value[Year %in% RefYears], na.rm = TRUE)

  SeriesList[[nm]] <- data.frame(Index = nm, Year = Year, Value = Value)
  TrendList[[nm]]  <- data.frame(
    Index = nm,
    Year  = c(min(FitYear), max(Year), ProjYear),
    Value = exp(predict(Fit, newdata = data.frame(FitYear = c(min(FitYear), max(Year), ProjYear))))
  )
  PointList[[nm]] <- data.frame(
    Index = nm,
    Year  = c(max(Year), ProjYear),
    Value = c(Status, IProj),
    Label = c('Status (most recent year)', 'Projected value (I_proj)')
  )
  IProjVec[nm] <- IProj
  IRefVec[nm]  <- IRef
}

SeriesDF <- dplyr::bind_rows(SeriesList)
TrendDF  <- dplyr::bind_rows(TrendList)
PointDF  <- dplyr::bind_rows(PointList)
PointDF$Label <- factor(PointDF$Label,
                         levels = c('Status (most recent year)', 'Projected value (I_proj)'))

Icur <- mean(IProjVec)
Iref <- mean(IRefVec)

TrendSolid  <- TrendDF |> dplyr::group_by(Index) |> dplyr::slice(1:2) |> dplyr::ungroup()
TrendDashed <- TrendDF |> dplyr::group_by(Index) |> dplyr::slice(2:3) |> dplyr::ungroup()

PlotFrom <- max(Year) - 15

ggplot() +
  geom_line(data = SeriesDF, aes(Year, Value), colour = 'grey50') +
  geom_point(data = SeriesDF, aes(Year, Value), colour = 'grey50') +
  geom_line(data = TrendSolid, aes(Year, Value), colour = '#0072B2', linewidth = 1.1) +
  geom_line(data = TrendDashed, aes(Year, Value), colour = '#0072B2', linewidth = 1.1,
            linetype = 'dashed') +
  geom_point(data = PointDF, aes(Year, Value, shape = Label, fill = Label),
             colour = '#D55E00', size = 3, stroke = 1.2) +
  scale_shape_manual(values = c(19, 21)) +
  scale_fill_manual(values = c('#D55E00', 'white')) +
  coord_cartesian(xlim = c(PlotFrom, max(TrendDF$Year) + 0.5)) +
  facet_wrap(~Index, scales = 'free_y') +
  labs(x = 'Year', y = 'Index value', shape = NULL, fill = NULL) +
  theme_bw() +
  theme(legend.position = 'bottom')
```

Averaging the three indices' projected values equally gives $I_{\text{cur}} =$
`r round(Icur, 3)`, and averaging their three 2007--2010 reference levels equally gives
$I_{\text{ref}} =$ `r round(Iref, 3)`, so $I_{\text{cur}}/I_{\text{ref}} =$
`r round(Icur / Iref, 3)` for this example based on the historical data.

Since each index's `Status` is in that index's own units, not absolute catch or biomass, it must be
calibrated before it can be used to set a catch limit. Each index is first turned into its own
catch-per-index rate, $\hat{r}_j$: the total catch across all fleets in the last historical year
(`CalibYears` = `r eval(formals(IndexRate_MP)$CalibYears)`)
divided by that particular index's own value in that same year:

$$
\hat{r}_j = \frac{\bar{C}_{\text{hist}}}{\bar{I}_{\text{hist},\,j}}
$$

where $j$ indexes the three indices, $\bar{C}_{\text{hist}}$ is the total catch in the last
historical year and $\bar{I}_{\text{hist},\,j}$ is index $j$'s own value in that same year.
Because $\hat{r}_j$ has units of catch *per unit of that index*, multiplying it by
$\text{Status}_j$ (index $j$'s own `Status`, in the index's own units) will give a catch value, once the
rate itself has been scaled and throttled as described next.

Before that final multiplication, each index's own rate $\hat{r}_j$ is scaled by the tuning
parameter, $m$, and throttled by a single response fraction, $\theta(I_{\text{cur}}/I_{\text{ref}})$
giving each index its own adjusted rate, $\text{AdjRate}_j$. Each index's own most recent value, $\text{Status}_j$, is then
multiplied by its own adjusted rate to give that index's own implied catch limit. The trial catch
limit is the equally-weighted average of the three:

$$
\text{AdjRate}_j = m \times \hat{r}_j \times \theta\!\left(\frac{I_{\text{cur}}}{I_{\text{ref}}}\right)
$$
$$
\text{TrialTAC} = \frac{1}{3}\sum_{j=1}^{3} \text{Status}_j \times \text{AdjRate}_j
$$

where $\text{Status}_j$ is index $j$'s own most recent-year value, $m$ is the tuning parameter, and
$\theta(x)$ ramps the response linearly between two control points:

$$
\theta(x) =
\begin{cases}
0 & x \le x_L \\[4pt]
\dfrac{x - x_L}{x_U - x_L} & x_L < x < x_U \\[4pt]
1 & x \ge x_U
\end{cases}
$$

with $x_L$ = `r eval(formals(IndexRate_MP)$HCRControlPointsIndex)[1]` and
$x_U$ = `r eval(formals(IndexRate_MP)$HCRControlPointsIndex)[2]`. When
$I_{\text{cur}}/I_{\text{ref}}$ is at or below $x_L$ the throttle zeroes out every index's implied
catch limit, at or above $x_U$ (i.e. at or above the historical reference level) the full
calibrated rate is applied, while in between the response ramps up linearly (@fig-hcr-indexrate).

```{r}
#| label: fig-hcr-indexrate
#| echo: false
#| fig-cap: "Index Rate response: the fraction of the calibrated catch rate applied, as a function of the current abundance index relative to its historical reference level."
plot_hcr(eval(formals(IndexRate_MP)$HCRControlPointsIndex),
         eval(formals(IndexRate_MP)$HCRControlPointsRate),
         eval(formals(IndexRate_MP)$RampType)[1],
         xlab = 'Current Index / Historical Reference Index',
         ylab = 'Fraction of Calibrated Catch Rate Applied')
```

The `Responsiveness` parameter, $R$, controls what fraction of the move from the previous cycle's catch
limit to `TrialTAC` is actually taken:

$$
\text{Mod} = \left(\frac{\text{TrialTAC}}{\text{TAC}_{\text{prev}}}\right)^{\!R}
$$
$$
\text{TAC} = \text{TAC}_{\text{prev}} \times \text{Mod}
$$

where $\text{TAC}_{\text{prev}}$ is the previous cycle's catch limit. The Index Rate variants used in the MSE have $R =$
`r eval(formals(IndexRate_MP)$Responsiveness)`, so a
proportional change in `TrialTAC` relative to $\text{TAC}_{\text{prev}}$ only moves the actual TAC
by that fraction on the log scale. 

Finally, this proportional change in the TAC is capped per management cycle; the two 
Index Rate variants (`IR`, `IR_20`) share this structure and differ only in the maximum TAC change cap (@tbl-cmp-summary).

### MCC {#sec-cmps-mcc}

Based on the Mostly Constant Catch (MCC) design ICCAT adopted for North Atlantic swordfish,
the MCC CMP uses the same abundance indices as Index Rate, together with the last historical year's
total landings, to set each management cycle's catch limit in three steps. 

Unlike Index Rate where the catch limit is iteratively adjusted from the previous cycle's TAC, MCC uses fixed 
baseline TAC that is stepped by a pre-agreed multiplier. There is no per-cycle absolute TAC change cap.

1. The indices reporting each year are combined into a single series, and that combined series'
   most recent value, $I_{\text{cur}}$, is compared with its own average over a historical
   reference period, $I_{\text{ref}}$, giving a ratio, $I_{\text{rat}}$.
2. $I_{\text{rat}}$ determines which of several pre-agreed bins applies (@tbl-mcc-steps), each with
   its own multiplier on the baseline catch level.
3. The catch limit is the baseline catch level -- the last historical year's total landings, scaled
   by the tuning parameter -- multiplied by that bin's multiplier.

Each of these steps is set out in full below.

MCC first combines the raw indices into a single equally-weighted mean combined index $I_y$
(@fig-mcc-combined):

$$
I_y = \frac{1}{n_y}\sum_{j} I_{j,\,y}
$$


```{r}
#| label: fig-mcc-combined
#| warning: false
#| echo: false
#| fig-width: 8
#| fig-height: 4.5
#| fig-cap: "The combined index from the historical data (grey), calculated as the equally-weighted mean of the individual indices, its average over the historical reference period (blue, $I_{\\text{ref}}$), and its most recent value (orange point, $I_{\\text{cur}}$)."
HistRefYears <- eval(formals(MCC_HCR)$HistRefYears)
RecentYears  <- eval(formals(MCC_HCR)$RecentYears)

ActiveCols <- which(!is.na(IndexValue[nrow(IndexValue), ]))
ActiveVal  <- IndexValue[, ActiveCols, drop = FALSE]

CombinedIndexEx <- rowMeans(ActiveVal, na.rm = TRUE)
CombinedDF <- data.frame(Year = Year, Value = CombinedIndexEx)

McuCur <- mean(utils::tail(CombinedIndexEx, RecentYears))
McuRef <- mean(CombinedIndexEx[Year %in% HistRefYears], na.rm = TRUE)
McuRat <- McuCur / McuRef

RefDF <- data.frame(Year = range(HistRefYears[HistRefYears %in% Year]), Value = McuRef)

ggplot(CombinedDF, aes(Year, Value)) +
  geom_line(colour = 'grey50') +
  geom_point(colour = 'grey50') +
  geom_line(data = RefDF, aes(Year, Value), colour = '#0072B2', linewidth = 1.3) +
  geom_point(data = data.frame(Year = max(Year), Value = McuCur),
             aes(Year, Value), colour = '#D55E00', size = 3) +
  labs(x = 'Year', y = 'Combined index value') +
  theme_bw()
```

 $I_{\text{cur}}$, the combined index's value in the most recent year (the orange point in @fig-mcc-combined),
is compared with $I_{\text{ref}}$, calculated as the average over a
`r paste(range(eval(formals(MCC_HCR)$HistRefYears)), collapse=' -- ')` historical reference period
(the blue line in @fig-mcc-combined):

$$
I_{\text{rat}} = \frac{I_{\text{cur}}}{I_{\text{ref}}}
$$

For this example using the historical data, $I_{\text{cur}} =$ `r round(McuCur, 3)`, $I_{\text{ref}} =$
`r round(McuRef, 3)`, so $I_{\text{rat}} =$ `r round(McuRat, 3)`.
$I_{\text{rat}}$ then determines which multiplier on the baseline catch level from the pre-specified bins applies (@tbl-mcc-steps).

```{r}
#| label: tbl-mcc-steps
#| tbl-cap: "The MCC step schedule showing the multipliers applied to the baseline TAC level based on the current index ratio."
#| echo: false
mcc_breaks <- eval(formals(MCC_HCR)$IratBreaks)
mcc_mult   <- eval(formals(MCC_HCR)$TACmult)

mcc_tbl <- data.frame(
  Ratio = c(
    sprintf('< %.2f', mcc_breaks[1]),
    sprintf('%.2f -- %.2f', mcc_breaks[-length(mcc_breaks)], mcc_breaks[-1]),
    sprintf('>= %.2f', mcc_breaks[length(mcc_breaks)])
  ),
  Multiplier = mcc_mult
)

knitr::kable(mcc_tbl, col.names = c('Combined Index / Historical Base ($I_{\\text{rat}}$)', 'TACbase multiplier'))
```

The catch limit is then the baseline catch level (the last historical year's total landings
across all fleets, scaled by the tuning parameter) multiplied by that bin's multiplier:

$$
\text{TAC}_{\text{base}} = C_{\text{LHY}} \times m
$$
$$
\text{TAC} = \text{TAC}_{\text{base}} \times s(I_{\text{rat}})
$$

where $C_{\text{LHY}}$ is the total landings across all fleets in the last historical year, $m$ is
the tuning parameter, and $s(\cdot)$ is the step function in @tbl-mcc-steps. As noted above, there
is no per-cycle change cap on top of this step (@tbl-cmp-summary).

### Ensemble {#sec-cmps-ensemble}

The Ensemble CMP combines MCC and Index Rate by averaging their two independently-computed TAC 
recommendations. Each cycle, MCC and Index Rate independently compute their own TAC recommendation
and the Ensemble's catch limit is the tuning-parameter-scaled average of the two:

$$
C_{\text{limit}} = m \times \text{mean}\big(C_{\text{MCC}},\ C_{\text{IndexRate}}\big)
$$

where $C_{\text{MCC}}$ and $C_{\text{IndexRate}}$ are MCC's and Index Rate's own recommendations,
each computed independently at its own individually-tuned setting for the blend (not necessarily
the same setting as the standalone MCC/Index Rate variants in @tbl-cmp-summary), and $m$ is the
Ensemble's own tuning parameter (@tbl-cmp-summary), applied on top of the blend.

Two variants are included: 

1. `Ens` blends MCC with the uncapped Index Rate component;
2. `Ens_20` blends MCC with the 20%-capped Index Rate component

The internal MCC/Index Rate tunepars and the Index Rate caps are shown in @tbl-cmp-summary. 


## Tuning Methodology {#sec-tuning-methodology}

Each CMP has a set of structural settings (which data it uses, how that data
is combined or smoothed, the shape of its response, how much its recommendation can change between
cycles) plus a single tuning parameter, `tunepar`, that scales how aggressively it converts that
data into a TAC recommendation. A two stage process is used to determine the optimal values for both 
the structural settings and `tunepar`:

1. Each structural setting is explored first, trying a small number of candidate values for that
   setting (e.g. `CalibYears`: 1 vs. 2 years; `Smooth`: on vs. off; `WeightMethod`: equal vs.
   inverse-CV vs. inverse-variance weighting). Once the optimal values for these settings are found,
   they are fixed and the process moves on to `tunepar`.
2. `tunepar` is then found by grid search over the full pooled Reference OM grid (i.e.,the nine
   OMs spanning the Growth x Natural Mortality uncertainty grid to find the highest-yield 
   value that still meets the Safety and Status objectives from Resolution 24-09 (@sec-performance-metrics),
   evaluated jointly across all nine OMs.

The resulting `tunepar` (and, where applicable, per-cycle change cap) for each CMP is
given in @tbl-cmp-summary. 

### Index Rate {#sec-tuning-indexrate}

For Index Rate (@sec-cmps-indexrate), step 1 fixed: 

- which indices to use (all three currently-reporting indices, `WeightMethod = 'equal'`); 
- the calibration window (`CalibYears` =`r eval(formals(IndexRate_MP)$CalibYears)`); 
- whether to smooth the index before use (`Smooth = FALSE`; LOESS smoothing is available in the CMP but did not improve
on the raw index); 
- the trend-projection window and horizon (`TrendYears` = `r eval(formals(IndexRate_MP)$TrendYears)`, `TrendHorizon` =
`r eval(formals(IndexRate_MP)$TrendHorizon)`); 
- the shape of the ramped throttle (`RampType = 'linear'`, control points shown in @fig-hcr-indexrate); and 
- the `Responsiveness` parameter (`r eval(formals(IndexRate_MP)$Responsiveness)`, @sec-cmps-indexrate). 

`tunepar` (the `IndexFactor` multiplier on the calibrated rate) was then set by the step-2 grid search, run
separately for the uncapped and 20%-capped per-cycle change cap, giving the two variants in
@tbl-cmp-summary.

### MCC {#sec-tuning-mcc}

For MCC (@sec-cmps-mcc), the historical reference window (`HistRefYears` =
`r paste(range(eval(formals(MCC_HCR)$HistRefYears)), collapse = '--')`), the step schedule itself
(`IratBreaks`/`TACmult`, @tbl-mcc-steps), and how the three indices are combined into one series
(`WeightMethod = 'equal'`) were fixed in step 1. 

`tunepar` (the multiplier on the reference catch level) was then set by the step-2 grid search, 
giving the single MCC variant in @tbl-cmp-summary.

### Ensemble {#sec-tuning-ensemble}

For Ensemble (@sec-cmps-ensemble), step 1 fixed how the two component CMPs (Index Rate and MCC) are combined
(`Combiner = 'mean'`). 

The two components' own tuning parameters (`MCC_tunepar`, `IR_tunepar`) were re-tuned jointly with the
Ensemble's own `tunepar` in step 2. The resulting values for both Ensemble variants are given in @tbl-cmp-summary.

