# Reference Operating Models {#sec-oms-reference}

```{r}
#| include: false
library(MSEtool)
library(ggplot2)
rnorm(5)
knitr::read_chunk("../analysis/03-OM-Plots.R")
```

```{r load_files}
#| include: false
```

The Reference Set OMs are currently designed as a factorial uncertainty grid with
the two axes of uncertainty (*Growth* & *Natural Mortality*) developed during the 
2026 Stock Assessment (see @sec-stock-assessment).

The *Growth* and *Natural Mortality* axes were both represented with three levels, 
corresponding to the 25th, 50th, and 75th percentiles of the estimated parameter distributions.
The 50th percentile level for each axis matches the Base Case stock assessment model (`S05`).  
The OM imported from this model is referred to as the *Reference OM*.

The three levels for each axis produces a 3 x 3 uncertainty grid with 9 reference operating models (@tbl-reference-oms). 
The von Bertalanffy growth parameters and the reference Lorenzen *M* value (at reference age 11)
at each percentile level are shown in @tbl-percentile-grid.

@fig-lifehistory shows the length- and natural mortality-at-age curves for each of the 9 reference grid OMs.

::: {.callout-note}
## Additional Reference Uncertainties
Additional operating models can be added to the Reference OMs, e.g., uncertainty 
in the steepness parameter of the Beverton-Holt stock recruit relationship or age/size of maturity.

This will be discussed by the SALB MSE Technical Team.
:::


::: {#tbl-reference-oms}
| # | OM | Growth | Natural Mortality |
|-- | ---|---|---|
| 1| G_25-M_25 | 25th  | 25th  |
| 2 | G_25-M_50 | 25th  | 50th  |
| 3 | G_25-M_75 | 25th  | 75th  |
| 4 | G_50-M_25 | 50th  | 25th  |
| 5 | G_50-M_50 | 50th  | 50th  |
| 6 | G_50-M_75 | 50th  | 75th  |
| 7 | G_75-M_25 | 75th  | 25th  |
| 8 | G_75-M_50 | 75th  | 50th  |
| 9 | G_75-M_75 | 75th  | 75th  |

Grid of nine reference operating models (OMs) spanning two axes of uncertainty, each with three levels.
:::

::: {#tbl-percentile-grid}
| Percentile | $L_\infty$ | $K$ | $t_0$  | $M$ |
|---|---|---|---| --- |
| 25th | 115.93 | 0.235 | -0.561 | 0.29 | 
| 50th (Base Case) | 121.24 | 0.238 | -0.891 | 0.36 |
| 75th | 125.02 | 0.237 | -1.217 | 0.44 |
von Bertalanffy growth parameters and reference natural mortality ($M$) at each level of the OM uncertainty grid. 
:::

```{r}
#| label: fig-lifehistory
#| echo: false
#| fig-cap: "Length-at-age and natural mortality-at-age, by operating model."
<<lifehistory>>
```

## Overview of Population Dynamics 

Each operating model was imported from the maximum likelihood estimates (MLE) of
the corresponding SS3 run, so within each OM the historical population dynamics 
are identical across simulations.

The figures below provide an overview of the population dynamics described in the nine Reference OMs:

- @fig-sb: historical spawning biomass  
- @fig-sbsb0: historical spawning biomass relative to equilibrium unfished level 
- @fig-sbsbmsy: historical spawning biomass relative to the MSY level  
- @fig-fmsy: historical fishing mortality relative to the $F_\text{MSY}$

@tbl-refpoints summarises the MSY reference points, along with terminal year $F/F_\text{MSY}$ and $SB/SB_\text{MSY}$, for each of the nine Reference OMs.

```{r}
#| label: fig-sb
#| echo: false
#| fig-cap: "Historical spawning biomass ($SB$), coloured by operating model and faceted by the axes of uncertainty"
<<ssb>>
```

```{r}
#| label: fig-sbsb0
#| echo: false
#| fig-cap: "Historical spawning biomass depletion ($SB/SB_0$), by operating model."
<<depletion>>
``` 

```{r}
#| label: fig-sbsbmsy
#| echo: false
#| fig-cap: 'Historical spawning biomass relative to MSY level ($SB/SB_{\text{MSY}}$), by operating model.'
<<depletion_msy>>
```

```{r}
#| label: fig-fmsy
#| echo: false
#| fig-cap: 'Historical fishing mortality relative to FMSY level ($F/F_{\text{MSY}}$), by operating model.'
<<F_msy>>
```

## Reference Points 
```{r}
#| label: tbl-refpoints
#| echo: false
#| tbl-cap: "MSY reference points and terminal year status, by operating model."
<<refpoints_table>>
```

 