---
suppress-bibliography: true
---

# Performance Metrics {#sec-performance-metrics}

```{r}
#| include: false
library(MSEtool)
library(ggplot2)
library(Slick)
OM <- readRDS("../objects/OM/Ref.om")
slick <- readRDS("../objects/slick.rds")
```

## Primary Performance Metrics

Resolution 24-09 [@ICCAT2024Res2409] describes the initial operational management
objectives for Southern Atlantic Albacore:

1. **Stock Status.** The stock should have a 60% or greater probability of occurring in the
green quadrant of the Kobe matrix over a 30-year projection period: i.e. $SB > SB_{MSY}$
and $F < F_{MSY}$ simultaneously.

2. **Safety.** There should be no greater than 15% probability of the stock falling below
$B_{LIM}$ at any point during the 30-year projection period, using an interim $B_{LIM}$ of
40% of $SB_{MSY}$. So
that higher values are better for every PM, this is expressed as the probability of the
stock remaining at or above $B_{LIM}$ throughout the projection period, with a target of
at least 85%.

3. **Yield.** The objective is to maximise overall catch levels. As the Resolution does not
specify a particular statistic, the mean annual catch is used. 
 
4. **Stability.** Any change in total allowable catch (TAC) between management periods
should be 20% or less, noting that asymmetric stability limits (e.g. different thresholds
for TAC increases versus decreases) may also be evaluated. This is expressed as the
probability that the change in TAC between successive management cycles is less than
20%; i.e. calculated between the TAC values actually set every `r Interval(OM)` years.

@tbl-pm-summary summarises these objectives together with the corresponding
performance metric and the associated target. The equation for each performance
metric is given below the table.

```{r}
#| echo: false
#| label: tbl-pm-summary
#| tbl-cap: "Summary of the Resolution 24-09 [@ICCAT2024Res2409] initial operational management objectives and the corresponding performance metrics used to evaluate CMP performance."
PrimarySummaryDF <- data.frame(
  Objective = c('Stock Status', 'Safety', 'Yield', 'Stability'),
  Metric    = c(
    'Probability of occurring in the Kobe green quadrant',
    'Probability of remaining at or above $B_{LIM}$ (40% $SB_{MSY}$) throughout the projection',
    'Mean annual catch',
    'Probability that the change in TAC between management periods is < 20%'
  ),
  Target    = c(
    '≥ 60%',
    '≥ 85%',
    'Maximise',
    'Ranked'
  )
)

knitr::kable(PrimarySummaryDF, col.names = c('Objective', 'Performance Metric', 'Target'))
```

Each PM can also be calculated for an individual OM or across a set of OMs (e.g. the Reference Set).
In the latter case, the simulations from every OM in the set are first combined 
into a single pooled MSE object and the summary statistic (e.g. a probability 
or a mean) is then calculated over that combined set of simulations. 

The performance metrics in @tbl-pm-summary are calculated as follows:

**Stock Status:**
$$
\text{Prob}\left(SB/SB_{MSY} > 1 \text{ and } F/F_{MSY} < 1\right)
$$

**Safety:**
$$
\text{Prob}\left(\min_y(SB_y/SB_{MSY}) \geq 0.4\right)
$$

**Yield:**
$$
\text{mean}_y(C_y)
$$

**Stability:**
$$
\text{Prob}\left(\lvert (TAC_c - TAC_{c-1})/TAC_{c-1}\rvert < 0.2\right)
$$

## Secondary Performance Metrics {#sec-secondary-pms}

Beyond the four primary metrics above, a wider set of performance metrics are reported for every
CMP/OM combination (@tbl-pm-secondary). 

Every performance metric (PM) is calculated over the full 30-year projection period. 
*Status*, *Yield*, *SB_SBMSY*, and *F_FMSY* are also broken down into three
time windows: 

1. Short: the first 10 years (2028 -- 2037);
2. Medium: the second 10 years (2038 -- 2047);
3. Long: the last 10 years (2048 -- 2057).
```{r}
#| echo: false
#| label: tbl-pm-secondary
#| tbl-cap: "Secondary performance metrics computed for every CMP/OM combination, in addition to the four primary metrics above."
PrimaryCodes <- c('Yield_All', 'Safety', 'Status_All', 'Stability')

bp <- slick@Boxplot
SecondaryDF <- data.frame(
  Code        = Slick::Code(bp),
  Label       = Slick::Label(bp),
  Description = Slick::Description(bp)
)
SecondaryDF <- SecondaryDF[!SecondaryDF$Code %in% PrimaryCodes, ]
rownames(SecondaryDF) <- NULL

knitr::kable(SecondaryDF, col.names = c('Code', 'Metric', 'Description'))
```




